\addcontentsline{toc}{chapter}{Abstract}
% Thesis Abstract -----------------------------------------------------
\chapter*{Abstract}
\label{ch:abstract}

%\begin{abstractslong}    %uncommenting this line, gives a different abstract heading
%\begin{abstracts}        %this creates the heading for the abstract page

\hspace*{0.3cm}
\indent With an increasing need to improve existing detector technology for both basic and applied nuclear science, an effort has been made to significantly further the development of two promising inorganic scintillators, Cs$_2$LiYCl$_6$ (CLYC) and CeBr$_3$. CLYC is a dual neutron/$\gamma$ scintillator with excellent pulse-shape discrimination and good energy resolution. Direct fast neutron spectroscopy is possible due to the $^{35}$Cl(n,p) reaction, whereas thermal neutron detection is accomplished using the $^{6}$Li(n,$\alpha$) reaction. A study of the fast neutron response in CLYC was performed at the UMass Lowell Van de Graaff using mono-energetic neutron beams between 0.8 to 2.6 MeV. Thermal neutron measurements were carried out using a moderated PuBe source and thermal beams from the UMass Research Reactor neutron radiography port. Simulations of the fast/thermal neutron response and efficiency were performed with the Monte Carlo code MCNPX (v2.7.0). Experiments with CeBr$_3$, a fast high light output $\gamma$-ray detector, focused on its excellent timing resolution ($\le$250 ps). A pair of CeBr$_3$ detectors were used for direct measurements of nanosecond and sub-nanosecond isomers in $^{152}$Sm and $^{177}$Hf using the delayed coincidence technique. The position resolution of a positron emitting source placed between detectors was evaluated to assess the potential use of CeBr$_3$ as a time-of-flight positron emission tomography (TOF PET) detector. Additional characterizations were performed for both CLYC and CeBr$_3$, including energy resolution, timing resolution, efficiency, and pulse shape analysis. 


%\end{abstracts}
%\end{abstractlongs}


% ---------------------------------------------------------------------- 